---
title: "Applied Linear Statistical Models"
author: "Ed Young"
date: "11/8/2019"
output:
     html_document:
          theme: null
          highlight: null
          css: styles.css
---
     
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, warning=FALSE, message=FALSE)
```
<!-- Load Libraries and Data -->
```{r check_packages, include=FALSE}
## required install on system (sudo apt install ...):  libgdal-dev, libudunits2-dev, libcurl4-openssl-dev
## List of packages to check and possibly install

packages <- c("tidyverse","broom","latex2exp","data.table","ggfortify","randtests","MASS","lmtest","here","moments","ALSM")
  
check_load_pkgs <- function(packages) {

  # Loop through each package to check if it is installed
  for (pkg in packages) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      message(paste("installing the '", pkg, "' package", sep = ""))
      
      # Error handling for installation
      tryCatch(
        {
          install.packages(pkg, Ncpus = parallel::detectCores())
        },
        error = function(e) {
          message(paste("Failed to install ", pkg, ": ", e, sep = ""))
        }
      )
    }
    # Attempt to load package
    load_status <- tryCatch({
      library(pkg, character.only = TRUE)
    }, error = function(e) {
      return(paste("Failed to load", pkg, "Error:", e))
    }) 
    
    # Check load status
    if (inherits(load_status, "try-error")) {
      message(load_status)
    } else {
      message(paste("Successfully loaded:", pkg))
    }
  }
}

# Run the function to check and install packages
check_load_pkgs(packages)
```

```{r load_data, include=FALSE}
df20<-read.csv("data/Chapter1DataSets/CH01PR20.txt", header=FALSE, sep=" ", col.names=c("Y","X"))
df21<-read.csv("data/Chapter1DataSets/CH01PR21.txt", header=FALSE, sep=" ", col.names=c("Y","X"))
df27<-read.csv("data/Chapter1DataSets/CH01PR27.txt", header=FALSE, sep=" ", col.names=c("Y","X"))
```

## Useful Formulas

A *family confidence coefficent* is the probability, in advance of sampling, that the entire family of statements (e.g. $\beta_0 and \beta_1$) will be within a specified limit.

****

### Woring-Hotelling Procedure

<span style="display:block" id="math">

</span>

****

### Bonferroni Procedure

<span style="display:block" id="math">
$s^2 =MSE = \frac{SSE}{n-2}=\frac{\sum\epsilon_i^2}{n-2}\\
s^2\{b_1\}=\frac{MSE}{\sum(X_i-\bar{X})^2}; s\{b_1\}=\sqrt{(\frac{MSE}{\sum(X_i-\bar{X})^2})}\\
s^2\{b_0\}=MSE\left [\frac{1}{n}+\frac{\bar{X}^2}{\sum(X_i-\bar{X})^2} \right ]\\
B=t(1-\frac{\alpha}{4}; n-2)\\
\textrm{Family Confidence Intervals: }\\
\quad b_0 \pm Bs\{b_0\}\\
\quad b_1 \pm Bs\{b_1\}\\\\$
<\span>

<span style="display:block" id="note">
Note that if $\bar{X}$ is positive, then $b_0$ and $b_1$ are negatively correlated - that they tend to err in opposite directions. 
<\span>

<span style="display:block" id="math">
$\sigma\{b_0, b_1\}=-\bar{X}\sigma^2\{b_1\}$
</span>

****

### Scheffé procedure

<span style="display:block" id="math">

</span>

****

### Berkson Model

<span style="display:block" id="math">

</span>

****

### Inverse Predictions

<span style="display:block" id="math">

</span>

****

# Chapter 4

## Problem 4.3

Refer to **Copier maintenance** Problem 1.20

### Will $b_0$ and $b_1$ tend to err in the same direction or in opposite directions here:  Explain.

To see whether $b_0$ and $b_1$ tend to err in the same direction or in opposite directions, I will leverage the fact that

$\sigma\{b_0, b_1\}=-\bar{X}\sigma^2\{b_1\}$

Because $\bar{X}=$ `r mean(df20$X)`, and is positive, it implies that $b_0$ and $b_1$ tend to err in opposite directions.

### Obtain Bonferroni joint confidence intervals for $\beta_0$ and $\beta_1$, using a 95% family confidence coefficient.

To obtain the Bonferroni joint confidence intervals, I will use the formulas:

<span style="display:block" id="math">
$s^2 =MSE = \frac{SSE}{n-2}=\frac{\sum\epsilon_i^2}{n-2}\\
s^2\{b_1\}=\frac{MSE}{\sum(X_i-\bar{X})^2}; s\{b_1\}=\sqrt{(\frac{MSE}{\sum(X_i-\bar{X})^2})}\\
s^2\{b_0\}=MSE\left [\frac{1}{n}+\frac{\bar{X}^2}{\sum(X_i-\bar{X})^2} \right ]\\
B=t(1-\frac{\alpha}{4}; n-2)\\
\textrm{Family Confidence Intervals: }\\
\quad b_0 \pm Bs\{b_0\}\\
\quad b_1 \pm Bs\{b_1\}$
</span>


```{r}
n<-length(df20$X)
alpha<-.05
model120<-lm(Y~X, data=df20)
b1<-model120$coefficients[2]
b0<-model120$coefficients[1]
MSE<-sum(model120$residuals^2)/(n-2)
s2b1<-MSE/sum((df20$X-mean(df20$X))^2)
sb1<-sqrt(s2b1)
s2b0<-MSE*((1/n)+(mean(df20$X)^2)/sum((df20$X-mean(df20$X))^2))
sb0<-sqrt(s2b0)
B<-qt(1-(alpha/4),n-2)
b1Lower<-b1-B*sb1
b1Upper<-b1+B*sb1
b0Lower<-b0-B*sb0
b0Upper<-b0+B*sb0
```

The 95% Family Confidence Intervals are:

`r b0Lower` $\leq \beta_0 \leq$ `r b0Upper`

`r b1Lower` $\leq \beta_1 \leq$ `r b1Upper`

### A consultant has suggested that $\beta_0$ should be 0 and $\beta_1$ should equal 14.0.  Do your joint confidence intervals in part (2) support his view?

Yes, those extimates are within the joint confidence intervals

## Problem 4.4

Refer to **Airfreight breakage** Problem 1.21

### Will $b_0$ and $b_1$ tend to err in the same direction or in opposite directions here?  Explain.

To see whether $b_0$ and $b_1$ tend to err in the same direction or in opposite directions, I will leverage the fact that

$\sigma\{b_0, b_1\}=-\bar{X}\sigma^2\{b_1\}$

Because $\bar{X}=$ `r mean(df21$X)`, and is positive, it implies that $b_0$ and $b_1$ tend to err in opposite directions.

### Obtain Bonferroni joint confidence intervals for $\beta_0$ and $\beta_1$, using a 99% family confidence coefficient.  Interpret your confidence intervals.

To obtain the Bonferroni joint confidence intervals, I will use the formulas:

<span style="display:block" id="math">
$s^2 =MSE = \frac{SSE}{n-2}=\frac{\sum\epsilon_i^2}{n-2}\\
s^2\{b_1\}=\frac{MSE}{\sum(X_i-\bar{X})^2}; s\{b_1\}=\sqrt{(\frac{MSE}{\sum(X_i-\bar{X})^2})}\\
s^2\{b_0\}=MSE\left [\frac{1}{n}+\frac{\bar{X}^2}{\sum(X_i-\bar{X})^2} \right ]\\
B=t(1-\frac{\alpha}{4}; n-2)\\
\textrm{Family Confidence Intervals: }\\
\quad b_0 \pm Bs\{b_0\}\\
\quad b_1 \pm Bs\{b_1\}$
</span>


```{r}
n<-length(df21$X)
alpha<-.01
model121<-lm(Y~X, data=df21)
b1<-model121$coefficients[2]
b0<-model121$coefficients[1]
MSE<-sum(model121$residuals^2)/(n-2)
s2b1<-MSE/sum((df21$X-mean(df21$X))^2)
sb1<-sqrt(s2b1)
s2b0<-MSE*((1/n)+(mean(df21$X)^2)/sum((df21$X-mean(df21$X))^2))
sb0<-sqrt(s2b0)
B<-qt(1-(alpha/4),n-2)
b1Lower<-b1-B*sb1
b1Upper<-b1+B*sb1
b0Lower<-b0-B*sb0
b0Upper<-b0+B*sb0
```

The 99% Family Confidence Intervals are:

`r b0Lower` $\leq \beta_0 \leq$ `r b0Upper`

`r b1Lower` $\leq \beta_1 \leq$ `r b1Upper`

## Problem 4.6

Refer to **Muscle mass** Problem 1.27

### Obtain Bonferroni joint confidence intervals for $\beta_0$ and $\beta_1$, using a 99% family confidence coefficient.  Interpret your confidence intervals.

To obtain the Bonferroni joint confidence intervals, I will use the formulas:

<span style="display:block" id="math">
$s^2 =MSE = \frac{SSE}{n-2}=\frac{\sum\epsilon_i^2}{n-2}\\
s^2\{b_1\}=\frac{MSE}{\sum(X_i-\bar{X})^2}; s\{b_1\}=\sqrt{(\frac{MSE}{\sum(X_i-\bar{X})^2})}\\
s^2\{b_0\}=MSE\left [\frac{1}{n}+\frac{\bar{X}^2}{\sum(X_i-\bar{X})^2} \right ]\\
B=t(1-\frac{\alpha}{4}; n-2)\\
\textrm{Family Confidence Intervals: }\\
\quad b_0 \pm Bs\{b_0\}\\
\quad b_1 \pm Bs\{b_1\}$
</span>


```{r}
n<-length(df27$X)
alpha<-.01
model127<-lm(Y~X, data=df27)
b1<-model127$coefficients[2]
b0<-model127$coefficients[1]
MSE<-sum(model127$residuals^2)/(n-2)
s2b1<-MSE/sum((df27$X-mean(df27$X))^2)
sb1<-sqrt(s2b1)
s2b0<-MSE*((1/n)+(mean(df27$X)^2)/sum((df27$X-mean(df27$X))^2))
sb0<-sqrt(s2b0)
B<-qt(1-(alpha/4),n-2)
b1Lower<-b1-B*sb1
b1Upper<-b1+B*sb1
b0Lower<-b0-B*sb0
b0Upper<-b0+B*sb0
```

The 99% Family Confidence Intervals are:

`r b0Lower` $\leq \beta_0 \leq$ `r b0Upper`

`r b1Lower` $\leq \beta_1 \leq$ `r b1Upper`



### Will $b_0$ and $b_1$ tend to err in the same direction or in opposite directions here?  Explain.

To see whether $b_0$ and $b_1$ tend to err in the same direction or in opposite directions, I will leverage the fact that

$\sigma\{b_0, b_1\}=-\bar{X}\sigma^2\{b_1\}$

Because $\bar{X}=$ `r mean(df27$X)`, and is positive, it implies that $b_0$ and $b_1$ tend to err in opposite directions.

### A researcher has suggested that $\beta_0$ should be 160 and $\beta_1$ should be between -1.9 and -1.5.  Do your joint confidence intervals in part (1) support his view?

No.  the researchers suggestion of $\beta_1$ is too low.

## Problem 4.7

Refer to **Copier maintenance** Problem 1.20

### Estimate the expected number of minutes spent when there are 3,5, and 7 copiers to be serviced, respectively.  Use interval estimates with a 90% family confidence coefficient based on the Working-Hotelling procedure.

### Two service calls for preventive maintenance are scheduled in which the numbers of copiers to be serviced are 4 and 7 respectively.  A family of prediction intervals for the times to be spent on these calls is desired with a 90% family confidence coefficient.  Which procedure, Scheffé or Bonferroni will provide tighter prediction limits here?

### Obtain the family of prediction intervals required in part (2), using the more efficient procedure.

## Problem 4.8

Refer to **Airfreight breakage** Problem 1.21

### It is desired to obtain interval estimates of the mean number of broken ampules when there are 0,1, and 2 transfers for a shipment, using a 95% family confidence coefficient.  Obtain the desired confidence intervals, using the Working-Hotelling procedure.

### Are the confidence intervals obtained in part (1) more efficient than Bonferroni intervals here?  Explain.

### The next three shipments will make 0,1, and 2 tranfers, respectively.  Obtain prediction intervals for the number of broken ampules for each of these three shipments, using the Scheffé procedure and a 95% family confidence coefficient.

### Would the Bonferroni procedure have been more efficient in developing the prediction intervals in part (3)?  Explain.

## Problem 4.10

Refer to **Muscle mass** Problem 1.27

### The nutritionist is

###

###

###

## Problem 4.16

Refer to **Copier maintenance** Problem 1.20

###

###

###


## Problem 4.17

Refer to **Copier maintenance** Problem 1.20

###

###

###


